---
title: "Lab3"
author: "Merlise Clyde"
date: "February 3, 2017"
output: 
   beamer_presentation:
          includes:
             in_header:  macros.tex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Topics

* wercker
* termplot
* more on interactions
* residuals in GLMS


##  wercker 

* Check your builds
* If there are missing packages let us know so that we may update the `werker.yml`
*  Should have passing badge at time of submission

## Termplot Example

```{r}
n = 100
logx1 = rnorm(n)
x1 = exp(logx1)
x2 = abs(rnorm(n))
logy = .5 + 3*logx1 - 2*sqrt(x2) + rnorm(n)
simdat = data.frame(X1=x1, X2 = x2, Y=exp(logy))
```

## Termplot output
```{r}
termplot(lm(Y ~ X1 + X2, data=simdat), terms = "X1",
         partial.resid = T, se=T, rug=T, smooth = panel.smooth)
```
## Termplot output
```{r}
termplot(lm(Y ~ X1 + X2, data=simdat), terms = "X2",
         partial.resid = T, se=T, rug=T, smooth = panel.smooth)
```

##  termplot with transformation of Y and X1

```{r}
termplot(lm(log(Y) ~ log(X1) + X2, data=simdat), terms = "log(X1)",
         partial.resid = T, se=T, rug=T, smooth = panel.smooth)
```
##  termplot with transformation of Y and X1

```{r}
termplot(lm(log(Y) ~ log(X1) + X2, data=simdat), terms = "X2",
         partial.resid = T, se=T, rug=T, smooth = panel.smooth)

```

##  termplot with transformation of Y, X1, and X2

```{r}
termplot(lm(log(Y) ~ log(X1) + sqrt(X2), data=simdat), terms = "sqrt(X2)",
         partial.resid = T, se=T, rug=T, smooth = panel.smooth)
```

## What is in a term plot?

*  x-axis is the (untransformed) variable in your dataframe  $(X1, X2)$
* line is the "term" of that variable's contribution to $f(x)$ 
* y-axis is partial residuals for term
* `partial.resid = T` adds the partial residuals to the plot
* `rug = T`  shows location of data on axes
* `se = T`  adds the SE of the term's contribution to $f(x)$
* `smooth = panel.smooth`  adds "smoothed" means to plot

##  Terms

$$Y = \hat{\beta}_0 + \hat{\beta}_1 X1 + \hat{\beta}_2 X2 + e$$
Equivalent to centered model

$$Y = \bar{Y} + \hat{\beta}_1 (X1 - \bar{X1}) + \hat{\beta}_2 (X2 - \bar{X2}) + e$$
Terms are coefficient estimates times centered predictors
 $$\hat{\beta}_1 (X1 - \bar{X1}) $$
 $$\hat{\beta}_2 (X2 - \bar{X2}) $$

##  Terms with transformations

$$\log(Y) = \hat{\beta}_0 + \hat{\beta}_1 \log(X1) + \hat{\beta}_2 X2 + e$$
Equivalent to centered model

$$\log(Y) = \bar{\log(Y)} + \hat{\beta}_1 (\log(X1) - \bar{\log(X1)}) + \hat{\beta}_2 (X2 - \bar{X2}) + e$$
Terms are coefficient estimates times centered "predictors"
 $$\hat{\beta}_1 (\log(X1) - \bar{\log(X1)}) $$

## partial residuals for a term

$$\log(Y) = \bar{\log(Y)} + \hat{\beta}_1 (\log(X1) - \bar{\log(X1)}) + \hat{\beta}_2 (X2 - \bar{X2}) + e$$

$$\log(Y) - (\bar{\log(Y)} + \hat{\beta}_1 (\log(X1) - \bar{\log(X1)})) =  \hat{\beta}_2 (X2 - \bar{X2}) + e$$

* Lefthand side takes response and removes the part of the response that is explained by $X1$  

* Equal to the `term` for $X2$ plus the residual $e$

* part of residual variation that is not explained by the other terms that potentially can be explained by $X2$  = partial residual for $X2$

* partial residual for X1  

$$\hat{\beta}_1 (\log(X1) - \bar{\log(X1)})  + e$$

##  All possible interactions

```{r}
lm(log(Y) ~ (log(X1) + sqrt(X2))^2, data=simdat)
```


## Selective proposal Interactions

Only add interactions if main effects are "significant"



